---
title: "Week_7_prac"
output: html_document
date: "2023-11-22"
---

## 1.introduction
```{r}
library(here)
library(janitor)
library(sf)
library(tidyverse)
```

```{r}
# read the ward data of London in

LondonWards <- st_read(here::here("Week_7_prac","London-wards-2018",
                                  "London-wards-2018_ESRI","London_Ward.shp"))

LondonWardsMerged <- st_read(here::here("Week_7_prac","London-wards-2018",
                                  "London-wards-2018_ESRI","London_Ward_CityMerged.shp")) %>%
  st_transform(., 27700)
```

```{r}
# read csv file in

WardData <- read_csv("https://data.london.gov.uk/download/ward-profiles-and-atlas/772d2d64-e8c6-46cb-86f9-e52b4c7851bc/ward-profiles-excel-version.csv",
                    locale = locale(encoding = "latin1"),
                     na = c("NA", "n/a")) %>% 
  clean_names()
```

```{r}
# join data and select the unique rows and select wanted columns

LondonWardsMerged <- LondonWardsMerged %>%
  left_join(., WardData,
            by = c("GSS_CODE" = "new_code")) %>%
  dplyr::distinct(., GSS_CODE, .keep_all = T) %>% # keep only unique rows, .keep_all if true, keep all variables in
  dplyr::select(., GSS_CODE, ward_name, average_gcse_capped_point_scores_2014)
```

```{r}
# check if crs of file is correct
st_crs(LondonWardsMerged)
```

```{r}
library(tmap)

BluePlaques <- st_read(here::here("Week_6_prac",
                                  "open-plaques-london-2018-04-08.geojson")) %>%
  st_transform(., 27700)
```

```{r}
# plot to check the data

tmap_mode("plot")

tm_shape(LondonWardsMerged) +
  tm_polygons(col = NA, alpha = 0.5) +
tm_shape(BluePlaques) +
  tm_dots(col = "blue")
```

## 2.data cleaning
```{r}
# we need to drop the points that fall outside the map

BluePlaques <- BluePlaques[LondonWardsMerged,]

# check the map again

tm_shape(LondonWardsMerged) +
  tm_polygons(col = NA, alpha = 0.5) +
tm_shape(BluePlaques) +
  tm_dots(col = "blue")
```

## 3.data manipulation
```{r}
example <- st_intersects(LondonWardsMerged, BluePlaques)

check_examples <- LondonWardsMerged %>%
  st_join(., BluePlaques) %>%
  filter(ward_name == "Kingston upon Thames - Coombe Hill")
```

```{r}
# take the length of each list per polygon and add this as new column

library(sf)

points_sf_joined <- LondonWardsMerged %>%
  mutate(n = lengths(st_intersects(., BluePlaques))) %>%
  janitor::clean_names() %>%
  # calculate the area
  mutate(area = st_area(.)) %>%
  # then density of the points per ward
  mutate(density = n/area) %>%
  # select density and some other variables
  dplyr::select(density, ward_name, gss_code, n, average_gcse_capped_point_scores_2014)
```

```{r}
# plot a quick choropleth map
points_sf_joined <- points_sf_joined %>%
  group_by(gss_code) %>%
  summarise(density = first(density),
            wardname = first(ward_name),
            plaquecount = first(n))

tm_shape(points_sf_joined) +
  tm_polygons("density", # 填充颜色根据变量density进行分级
              style = "jenks", # 分级方式为Jenks自然断点分级法
              palette = "PuOr", # 调色板
              midpoint = NA, # 地图中点的设置为默认值
              popup.vars = c("wardname", "density"), # 弹出信息框显示的内容
              title = "Blue Plaque Density")
```

## 4.weight matrix
```{r}
library(spdep)
```
```{r}
# first calculate the centroids of all wards in London
coordsW <- points_sf_joined %>%
  st_centroid(.) %>% # 提取中心坐标
  st_geometry(.) # 将中心创建为一个新的sf对象

plot(coordsW, axes = TRUE) # axes = TRUE 坐标轴将被包含在绘图中
```

```{r}
# create a neighbours list
LWard_nb <- points_sf_joined %>%
  poly2nb(., queen=T)
```
```{r}
summary(LWard_nb)
```
```{r}
# now we can plot the neighbours
plot(LWard_nb, st_geometry(coordsW), col = "red")
# add a map underneath
plot(points_sf_joined$geometry, add = T)
```
matrix style
```{r}
# create a spatial weights matrix from these weights
Lward.lw <- LWard_nb %>%
  nb2mat(., style="B") # B为二进制，近表示是否相邻，不考虑邻接关系的强弱

sum(Lward.lw) # 计算矩阵总和

sum(Lward.lw[1,]) # 计算矩阵第一行和
```

## 5.autocorrelation
```{r}
# we can do Moran’s I
# Moran’s I requires a spatial weight list type object
Lward.lw <- LWard_nb %>%
  nb2listw(., style="C")
```
Moran’s I
```{r}
I_LWard_Global_Density <- points_sf_joined %>%
  pull(density) %>%
  as.vector() %>%
  moran.test(., Lward.lw)

I_LWard_Global_Density
```
Geary’s C
```{r}
C_LWard_Global_Density <- points_sf_joined %>%
  pull(density) %>%
  as.vector() %>%
  geary.test(., Lward.lw)

C_LWard_Global_Density
```
Getis Ord
```{r}
G_LWard_Global_Density <- points_sf_joined %>%
  pull(density) %>%
  as.vector() %>%
  globalG.test(., Lward.lw)

G_LWard_Global_Density
```
So the global statistics are indicating that we have spatial autocorrelation of Blue Plaques in London

Local Moran’s I
```{r}
# use the localmoran function to generate I for each ward in the city
I_LWard_Local_count <- points_sf_joined %>%
  pull(plaquecount) %>% # in local Moran's I pull use different attribute
  as.vector() %>%
  localmoran(., Lward.lw) %>%
  as_tibble()

I_LWard_Local_Density <- points_sf_joined %>%
  pull(density) %>%
  as.vector() %>%
  localmoran(., Lward.lw) %>%
  as_tibble()

# and we can choose the first 5 rows
slice_head(I_LWard_Local_Density, n=5)
```
There are 5 columns of data. We want to copy some of the columns (the I score (column 1) and the z-score standard deviation (column 4)) back into the LondonWards spatialPolygonsDataframe
```{r}
points_sf_joined <- points_sf_joined %>%
  mutate(plaque_count_I = as.numeric(I_LWard_Local_count$Ii)) %>%
  mutate(plaque_count_Iz =as.numeric(I_LWard_Local_count$Z.Ii)) %>%
  mutate(density_I =as.numeric(I_LWard_Local_Density$Ii)) %>%
  mutate(density_Iz =as.numeric(I_LWard_Local_Density$Z.Ii))
```
Mapping
```{r}
# set the breaks manually
breaks1<-c(-1000,-2.58,-1.96,-1.65,1.65,1.96,2.58,1000)
```
```{r}
library(RColorBrewer)
MoranColours<- rev(brewer.pal(8, "RdGy"))
```
plot
```{r}
tm_shape(points_sf_joined) +
  tm_polygons("plaque_count_Iz",
              style = "fixed", # 使用固定样式，地图上的颜色样式基于具体的数值
              breaks = breaks1,
              palette = MoranColours,
              midpoint = NA,
              title = "Local Moran's I, Blue Plaques in London")
```

Local Getis Ord G
```{r}
Gi_LWard_Local_Density <- points_sf_joined %>%
  pull(density) %>%
  as.vector() %>%
  localG(., Lward.lw)

head(Gi_LWard_Local_Density)
```
```{r}
# add the local G to sf
points_sf_joined <- points_sf_joined %>%
  mutate(density_G = as.numeric(Gi_LWard_Local_Density))
```
plot
```{r}
library(RColorBrewer)

GIColours<- rev(brewer.pal(8, "RdBu"))

#now plot on an interactive map
tm_shape(points_sf_joined) +
    tm_polygons("density_G",
        style="fixed",
        breaks=breaks1,
        palette=GIColours,
        midpoint=NA,
        title="Gi*, Blue Plaques in London")
```

## 6.other variables
```{r}
# use head to see what other variables are in the data file
slice_head(points_sf_joined, n=2)
```
```{r}
Datatypelist <- LondonWardsMerged %>% 
  st_drop_geometry()%>%
  summarise_all(class) %>%
  pivot_longer(everything(), 
               names_to="All_variables", 
               values_to="Variable_class")

Datatypelist
```
```{r}
LondonWardsMerged$average_gcse_capped_point_scores_2014 <- ifelse(is.na(LondonWardsMerged$average_gcse_capped_point_scores_2014), 0, LondonWardsMerged$average_gcse_capped_point_scores_2014)

I_LWard_Local_GCSE <- LondonWardsMerged %>%
  arrange(GSS_CODE) %>% # 对数据框进行排序
  pull(average_gcse_capped_point_scores_2014) %>%
  as.vector() %>%
  localmoran(., Lward.lw) %>%
  as_tibble()

points_sf_joined <- points_sf_joined %>%
  arrange(gss_code)%>%
  mutate(GCSE_LocIz = as.numeric(I_LWard_Local_GCSE$Z.Ii))

tm_shape(points_sf_joined) +
    tm_polygons("GCSE_LocIz",
        style="fixed",
        breaks=breaks1,
        palette=MoranColours,
        midpoint=NA,
        title="Local Moran's I, GCSE Scores")
```

```{r}
G_LWard_Local_GCSE <- LondonWardsMerged %>%
  dplyr::arrange(GSS_CODE)%>%
  dplyr::pull(average_gcse_capped_point_scores_2014) %>%
  as.vector()%>%
  localG(., Lward.lw)

points_sf_joined <- points_sf_joined %>%
  dplyr::arrange(gss_code)%>%
  dplyr::mutate(GCSE_LocGiz = as.numeric(G_LWard_Local_GCSE))

tm_shape(points_sf_joined) +
    tm_polygons("GCSE_LocGiz",
        style="fixed",
        breaks=breaks1,
        palette=GIColours,
        midpoint=NA,
        title="Gi*, GCSE Scores")
```

