---
title: "Week_2_homework"
output: html_document
date: "2023-10-12"
---

# HOMEWORK!

## 1. load library

```{r}
library(readr)
library(here)
library(sf)
library(tmap)
library(tmaptools)
library(tidyverse)
library(janitor)
library(OpenStreetMap)
```

## 2. loda data

### 2.1 load map data

```{r}
WC <- st_read(here::here("Week_2_prac",
                         "homework",
                         "Washington_Counties_with_Natural_Shoreline_washsh_area",
                         "Washington_Counties_with_Natural_Shoreline_washsh_area.shp"))
```

### 2.2 check the shapefile

```{r}
qtm(WC)
```

### 2.3 load csv

```{r}
original_data <- read_csv(here::here("Week_2_prac",
                                     "homework",
                            "Report_Card_Assessment_Data_2018-19_School_Year_20231012.csv"),
                          na="NULL")
```

### 2.4 clean data

#### 2.4.1 clean names and select appropriate rows

```{r}
school_data <- original_data %>%
  clean_names()
school_data <- school_data %>%
  filter(grade_level == "All Grades") %>%
  filter(student_group_type == "All") %>%
  filter(test_subject == "Science")
```

#### 2.4.2 select appropriate columns

```{r}
school_data1 <- school_data[c(3,20,22)]
```

#### 2.4.3 replace "NULL" with "0"

```{r}
school_data1$count_of_students_expected_to_test[school_data1$count_of_students_expected_to_test == "NULL"] <- 0
school_data1$count_met_standard[school_data1$count_met_standard == "NULL"] <- 0
```
replace "na" with "0"
```{r}
school_data1[is.na(school_data1)] <- 0
```

#### 2.4.4 transform character to numeric

```{r}
school_data1$count_of_students_expected_to_test <- as.numeric(school_data1$count_of_students_expected_to_test)
school_data1$count_met_standard <- as.numeric(school_data1$count_met_standard)
```
another way to transform character to numeric
```{r}
school_data1 <- school_data1 %>%
  mutate(count_of_students_expected_to_test=(as.numeric(count_of_students_expected_to_test))) %>%
  mutate(count_met_standard=(as.numeric(count_met_standard)))
```


#### 2.4.5 aggregate rows and sum

```{r}
school_data2 <- aggregate(.~county,data = school_data1,sum)
```
another way to sum
```{r}
school_data2 <- school_data1 %>%
  group_by(county) %>%
  summarise(students_expected_to_test=sum(count_of_students_expected_to_test),
            count_met_standard=sum(count_met_standard))
```

#### 2.4.6 calculate the average of the whole state

```{r}
sum_col <- colSums(school_data2[,-1])
sum_col
state_average <- round((sum_col[2]/sum_col[1]),digits = 3)
state_average
```

### 2.5 calculate the percentage

```{r}
school_data2 <- school_data2 %>%
  mutate(average_percent=count_met_standard/count_of_students_expected_to_test) %>%
  mutate(across(where(is.numeric),round,3))
```

### 2.6 calculate the difference

```{r}
school_data2 <- school_data2 %>%
  mutate(difference = average_percent - state_average)
```

## 3. merge data

```{r}
b <- toupper(school_data2$county)
school_data3 <- school_data2
school_data3$county <- b
```
```{r}
BoroughDatamap <- WC %>%
  clean_names() %>%
  merge(.,school_data3,
        by.x="county",
        by.y="county")
```

## 4. plot
```{r}
tmap_mode("plot")
qtm(BoroughDatamap,fill = "difference")
```

### 4.1 create osm
```{r}
tmapwc <- BoroughDatamap %>%
  st_bbox(.) %>%
  tmaptools::read_osm(.,type = "osm",zoom=NULL)
```

### 4.2 plotting

```{r}
tmap_mode("plot")
tm_shape(tmapwc) +
tm_rgb() +
tm_shape(BoroughDatamap) +
  tm_polygons("difference",
              midpoint = 0,
              title = "County average compared with the state average",
              alpha = 0.5) +
  tm_compass(position = c("left","bottom"),type = "arrow") +
  tm_scale_bar(position = c("left","bottom")) +
  tm_layout(title = "Difference in students who met standards",legend.position = c("right", "bottom"))
```

