---
title: "Week_3_prac"
output: html_document
date: "2023-10-17"
---

# Week3!

## 1. load data

### 1.1 load library

```{r}
library(sf)
library(here)
```

### 1.2 load geopackage

```{r}
st_layers(here::here("Week_3_prac","gadm36_AUS.gpkg"))
```

```{r}
Ausoutline <- st_read(here::here("Week_3_prac","gadm36_AUS.gpkg"),layer = "gadm36_AUS_0")
```

```{r}
print(Ausoutline)
```

### 1.3 proj4

```{r}
library(sf)
st_crs(Ausoutline)$proj4string
```

### 1.4 EPSG

```{r}
Ausoutline <- Ausoutline %>%
  st_set_crs(.,4326) # if data didn't have a crs, we can use this one
```

### 1.5 reprojecting spatial data

```{r}
AusoutlineProjected <- Ausoutline %>%
  st_transform(.,3112)
print(AusoutlineProjected)
```

```{r}
# transform from sf to sp
# AusoutlineSP <- Ausoutline %>%
#   as(.,"Spatial")

# transform from sp to sf
# AusoutlineSF <- AusoutlineSP %>%
#   st_as_sf()

#sp包是传统处理spatial的包，现主要用sf包
```

## 2. worldclimate data

### 2.1 load library and load climate data

```{r}
library(raster)
library(terra)
jan <- terra::rast(here::here("Week_3_prac","wc2.1_5m_tavg","wc2.1_5m_tavg_01.tif"))
jan
plot(jan)
```

### 2.2 transform the crs of a raster

one solution

```{r}
# pr1 <- terra::project(jan,
#                       "+proj=moll +lon_0=0 +x_0=0 +y_0=0 +ellps=WGS84 +datum=WGS84 +units=m +no_defs")
# plot(pr1)
```

another easy solution

```{r}
newproj <- 'ESRI:54009'
pr1 <- jan %>%
  terra::project(.,newproj)
plot(pr1)
```

to WGS84

```{r}
pr1 <- pr1 %>%
  terra::project(.,"EPSG:4326")
plot(pr1)
```

## 3. data loading

### 3.1 dir_info

```{r}
library(fs)
dir_info('Week_3_prac') # illustrate what is in the directory
dir_info('Week_3_prac/wc2.1_5m_tavg')
```

### 3.2 select data

```{r}
library(tidyverse)
listfiles <- dir_info('Week_3_prac/wc2.1_5m_tavg') %>%
  filter(str_detect(path,'.tif')) %>%
  dplyr::select(path) %>%
  pull()
```

### 3.3 load raster

```{r}
worldclimtemp <- listfiles %>%
  terra::rast()
worldclimtemp
```

### 3.4 access the single layer in SpatRaster

```{r}
worldclimtemp[[1]]
```

### 3.5 rename the name of each layer

```{r}
month <- c("Jan", "Feb", "Mar", "Apr", "May", "Jun", 
           "Jul", "Aug", "Sep", "Oct", "Nov", "Dec")
names(worldclimtemp) <- month
worldclimtemp[[1]]
worldclimtemp$Mar
```

### 3.6 raster location

```{r}
site <- c("Brisbane", "Melbourne", "Perth", "Sydney", "Broome", "Darwin", "Orange", 
          "Bunbury", "Cairns", "Adelaide", "Gold Coast", "Canberra", "Newcastle", 
          "Wollongong", "Logan City")
lon <- c(153.03, 144.96, 115.86, 151.21, 122.23, 130.84, 149.10, 115.64, 145.77, 
         138.6, 153.43, 149.13, 151.78, 150.89, 153.12)
lat <- c(-27.47, -37.91, -31.95, -33.87, 17.96, -12.46, -33.28, -33.33, -16.92, 
         -34.93, -28, -35.28, -32.93, -34.42, -27.64)

# put all of this information into one list
samples <- data.frame(site,lon,lat,row.names = 'site') # why?

# extract the data from the Rasterstack
AUcitytemp <- terra::extract(worldclimtemp,samples)
```

### 3.7 add city names

```{r}
Aucitytemp2 <- AUcitytemp %>%
  as_tibble() %>%
  add_column(site=site,.before = 'Jan')
```

## 4. descriptive statistics

### 4.1 data preparation
take Perth as an example
```{r}
Perthtemp <- Aucitytemp2 %>%
  filter(site=='Perth')

# another way
Perthtemp <- Aucitytemp2[3,]
```

### 4.2 histogram
```{r}
hist(as.numeric(Perthtemp)) # tibble stored the data as double
                            # hist() function needs numeric
```
### 4.3 remove outlier and improve the aesthetics
```{r}
library(tidyverse)
# define the breaks in the historgram
userbreak <- c(8,10,12,14,16,18,20,22,24,26)
# remove id and site cols
t <- Perthtemp %>%
  dplyr::select(Jan:Dec) # dplyr::select(3:14) this is ok either
hist((as.numeric(t)),
     breaks=userbreak,
     col='red',
     main='Histogram of Perth Temperature',
     xlab='Temperature',
     ylab='Frequency')
```
```{r}
histinfo <- as.numeric(t) %>%
  as.numeric() %>% # why we need the code in this row?
  hist(.)
histinfo
```
### 4.4 using more data
```{r}
plot(Ausoutline$geom)
```
```{r}
AusoutSIMPLE <- Ausoutline %>% # simplify the outline, just for checking the geometry
  st_simplify(.,dTolerance = 1000) %>%
  st_geometry() %>%
  plot()
```
```{r}
print(Ausoutline) # check coordinate reference system
crs(worldclimtemp)
```
```{r}
Austemp <- worldclimtemp %>%
  terra::crop(.,Ausoutline) %>% # 把遥感数据裁剪到底图大小
  terra::mask(.,Ausoutline) # 把遥感数据底图中多余的部分遮挡起来
plot(Austemp)
```
```{r}
hist(Austemp$Mar,col='red',main='March temperature')
```
### 4.5 histogram with ggplot
```{r}
Ausdf <- Austemp %>% # make the raster into a data.frame
  as.data.frame()

library(ggplot2)
# set up the basic histogram
gghist <- ggplot(Ausdf,aes(x=Mar))+ # aes()用于定义映射关系
  geom_histogram(color='white',fill='black')+
  labs(title = "Ggplot2 histogram of Australian March temperatures",
                        x="Temperature",
                        y="Frequency")
# add a vertical line to the histogram showing mean tempearture
gghist + geom_vline(aes(xintercept=mean(Mar,
                                        na.rm=TRUE)),
                    color='red',
                    linetype='dashed',
                    size=1) +
  theme(plot.title = element_text(hjust = 0.5)) # 控制标题居中设置的
```
### 4.6 plotting multiple months
```{r}
squishdata <- Ausdf %>% # pivot_longer
  pivot_longer(cols = 1:12,
               names_to = 'Month',
               values_to = 'Temp')
```
```{r}
twomonths <- squishdata %>% # select two months
  filter(.,Month=='Jan'|Month=='Jun')
```
```{r}
meantwomonths <- twomonths %>%
  group_by(Month) %>%
  summarise(mean=mean(Temp,na.rm=TRUE))
meantwomonths
```
```{r}
ggplot(twomonths,aes(x=Temp,color=Month,fill=Month)) +
  geom_histogram(position = 'identity',alpha=0.5) +
  geom_vline(data=meantwomonths,
             aes(xintercept=mean,
                 color=Month),
             linetype='dashed') +
  labs(title = 'Ggplot2 histogram of Australian Jan and Jun temperatures',
       x='Temperature',
       y='Frequency') +
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5))
```
```{r}
data_complete_cases <- squishdata %>%
  drop_na() %>% # drop rows where any column contains a missing value
  mutate(Month=factor(Month,levels=c("Jan","Feb","Mar",
                                          "Apr","May","Jun",
                                          "Jul","Aug","Sep",
                                          "Oct","Nov","Dec")))
```
```{r}
ggplot(data_complete_cases,aes(x=Temp,na.rm=TRUE)) +
  geom_histogram(color="black",binwidth = 5) +
  labs(title="Ggplot2 faceted histogram of Australian temperatures",
       x="Temperature",
       y="Frequency") +
  facet_grid(Month~.) + # 把十二个月分开来画图
  theme(plot.title = element_text(hjust = 0.5))
```
### 4.7 interactive histogram using plotly
```{r}
library(plotly)

# split the data for plotly based on month

jan <- squishdata %>%
  drop_na() %>%
  filter(.,Month=='Jan')

jun <- squishdata %>%
  drop_na() %>%
  filter(.,Month=='Jun')

# give axis titles

x <- list(title = "Temperature")
y <- list(title = "Frequency")

# set the bin width

xbinsno <- list(start=0, end=40, size = 2.5)

# plot the histogram calling all the variables we just set

ihist<-plot_ly(alpha = 0.6) %>%
        add_histogram(x = jan$Temp,
        xbins=xbinsno, name="January") %>%
        add_histogram(x = jun$Temp,
        xbins=xbinsno, name="June") %>% 
        layout(barmode = "overlay", xaxis=x, yaxis=y)

ihist
```

### 4.8 other
```{r}
# mean per month
meanofall <- squishdata %>%
  group_by(Month) %>%
  summarise(mean=mean(Temp,na.rm=TRUE))

head(meanofall,n=10)
```
```{r}
# standard deviation per month
sdofall <- squishdata %>%
  group_by(Month) %>%
  summarise(sd=sd(Temp,na.rm=TRUE))
print(sdofall)
```
```{r}
# maximum per month
maxofall <- squishdata %>%
  group_by(Month) %>%
  summarise(max=max(Temp,na.rm=TRUE))
print(maxofall)
```
```{r}
# minimum per month
minofall <- squishdata %>%
  group_by(Month) %>%
  summarise(min=min(Temp,na.rm=TRUE))
print(minofall)
```
```{r}
# Interquartlie range per month
IQRofall <- squishdata %>%
  group_by(Month) %>%
  summarize(IQR = IQR(Temp, na.rm=TRUE))
print(IQRofall)
```
```{r}
# perhaps you want to store multiple outputs in one list..
lotsofstats <- squishdata %>%
  group_by(Month) %>%
  summarize(IQR = IQR(Temp, na.rm=TRUE), 
            max=max(Temp, na.rm=T))
```
```{r}
# or you want to know the mean (or some other stat) 
#for the whole year as opposed to each month...
meanwholeyear=squishdata %>%
  summarize(meanyear = mean(Temp, na.rm=TRUE))
```


