---
title: "Week_2_prac"
output: html_document
date: "2023-10-11"
---

# Week 2 prac

## load library

```{r}
library(sf)
library(tidyverse)
library(dplyr)
library(janitor)
library(plotly)
library(maptools)
```

## introduction

```{r}
mycsv <- read_csv("Week_1_prac/fly-tipping-borough.csv")
```

### show list

```{r}
ls()
```

### remove object

```{r}
rm(mycsv)
```

### basic plotting

```{r}
# create some datasets, first a vector of 1-100 and 101-200
Data1 <- c(1:100)
Data2 <- c(101:200)
# plot the data
plot(Data1,Data2,col="red")
```

### for help

```{r}
?plot
```

### data structures

```{r}
df <- data.frame(Data1,Data2)
plot(df,col="green")
df %>% 
  head(10)
df %>%
  tail(10)
```

### elements of a data frame

```{r}
df[2,1]
df[2,2]
df[1:10,2]
df[5:15,]
df[c(2,3,6),]
df[,1]
```

### rename

```{r}
df <- df %>%
  dplyr::rename(column1=Data1,column2=Data2)
```

```{r}
df %>%
  dplyr::select(column1)
```

## reading data

### load data

```{r}
LondonDataOSK <- read.csv("Week_2_prac/LondonData.csv",
                          header=TRUE,
                          sep=",",
                          encoding="latin1")
```

### another way to read files

```{r}
install.packages("here")
library(here)
```

```{r}
```

```{r}
# LondonDataOSK <- read.csv(here::here("Week_2_prac","LondonData.csv",header=TRUE,sep=",",encoding="latin1"))
# 无法使用，需请教
```

### read data from the web directly

```{r}
LondonData <- read_csv("https://data.london.gov.uk/download/ward-profiles-and-atlas/772d2d64-e8c6-46cb-86f9-e52b4c7851bc/ward-profiles-excel-version.csv",
                       locale = locale(encoding = "latin1"),
                       na="n/a")
```

### examine data

```{r}
class(LondonData)
class(LondonDataOSK)
```

```{r}
Datatypelist <- LondonData %>%
  summarise_all(class) %>%
  pivot_longer(everything(),
               names_to = "All_variables",
               values_to = "Variable_class")
Datatypelist
```

quickly edit data

```{r}
LondonData <- edit(LondonData)
```

quickly summarise data

```{r}
summary(df)
LondonData %>%
  colnames() %>%
  head(5)
```

### data manipulation

```{r}

```

#### select rows

```{r}
LondonBoroughs <- LondonData[626:658,]
```

#### filter工具

```{r}
Femalelifeexp <- LondonData %>%
  filter(.,`Female life expectancy -2009-13`>90) # why we should use the symbol ` ?
LondonBoroughs <- LondonData %>%
  filter(str_detect(`New code`,"^E09")) # ^加不加这个符号结果都一致？
```

#### check

```{r}
LondonBoroughs$`Ward name`
```

#### another way to check

```{r}
LondonBoroughs %>%
  dplyr::select(`Ward name`) %>%
  print()
```

#### 去重

```{r}
LondonBoroughs <- LondonBoroughs %>%
  distinct()
```

#### select columns

```{r}
LondonBoroughs_dplyrcols <- LondonBoroughs %>%
  dplyr::select(c(1,19,20,21))
```

#### another way to select columns

```{r}
LondonBoroughs_manualcols <- LondonBoroughs[,c(1,19,20,21)] # another way to select columns
```

#### the third way to select columns

```{r}
LondonBoroughs_contains <- LondonBoroughs %>%
  dplyr::select(contains("expectancy"),contains("obese - 2011/12 to 2013/14"),contains("Ward name"))
```
#### rename columns
```{r}
LondonBoroughs <- LondonBoroughs %>%
  dplyr::rename(Borough="Ward name") %>%
  clean_names()
```
#### calculate the average and select columns and arrange in descending order
```{r}
Life_expectancy <- LondonBoroughs %>%
  mutate(averagelifeexpectancy=((female_life_expectancy_2009_13+male_life_expectancy_2009_13)/2)) %>%
  mutate(normalisedlifeexpectancy=averagelifeexpectancy/mean(averagelifeexpectancy)) %>%
  dplyr::select(new_code,
                borough,
                averagelifeexpectancy,
                normalisedlifeexpectancy) %>%
  arrange(desc(normalisedlifeexpectancy))
```
#### compare the expectancy of London with that of UK
```{r}
Life_expectancy2 <- Life_expectancy %>%
  mutate(UKcompare=case_when(averagelifeexpectancy>81.16~"above UK average",
                             TRUE~"below UK average"))
```
```{r}
Life_expectancy2_group <- Life_expectancy2 %>%
  mutate(UKdiff=averagelifeexpectancy-81.16) %>%
  group_by(UKcompare) %>%
  summarise(range=max(UKdiff)-min(UKdiff),count=n(),average=mean(UKdiff))
Life_expectancy2_group
```
#### 头大的步骤
```{r}
Life_expectancy3 <- Life_expectancy %>%
  mutate(UKdiff = averagelifeexpectancy-81.16) %>%
  mutate(across(where(is.numeric),round,3)) %>%
  mutate(across(UKdiff,round,0)) %>%
  mutate(UKcompare=case_when(averagelifeexpectancy >= 81~
                               str_c("equal or above UK average by",
                                     UKdiff,"years",sep = " "),
                             TRUE~str_c("below UK average by",
                                        UKdiff,"years",sep = " "))) %>%
  group_by(UKcompare) %>%
  summarise(count=n())
Life_expectancy3
```
接着头大
```{r}
Life_expectancy4 <- Life_expectancy %>%
  mutate(UKdiff=averagelifeexpectancy-81.16) %>%
  mutate(across(where(is.numeric),round,3)) %>%
  mutate(across(UKdiff,round,0))
```
### plotting
```{r}
plot(LondonBoroughs$male_life_expectancy_2009_13,
     LondonBoroughs$percent_children_in_reception_year_who_are_obese_2011_12_to_2013_14)
```
#### pimp my graph
```{r}
plot_ly(LondonBoroughs,
        x= ~male_life_expectancy_2009_13,
        y= ~percent_children_in_reception_year_who_are_obese_2011_12_to_2013_14,
        text= ~borough,
        type="scatter",
        mode="markers")
```


